---
title: "all models results"
author: "Eren Kafadar"
date: "2025-06-23"
output: html_document
editor_options: 
  chunk_output_type: console
---
#Main Setup
```{r setup, include=FALSE}
save_tables <- FALSE
save_figures <- FALSE
code_path <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/longitudinal-models-abcd/"
save_path <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/3.0_results/"
data_path <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/CSV/process_tables_5.1_2.0/update/"
knitr::opts_chunk$set(echo = TRUE)
#Load libraries
library(dplyr)
library(magrittr)
library(readr)
library(stringr)
library(gamlss) #to fit model
require(gamlss.ggplots)
library(splines)
library(tidyr)
library(gamlssTools) #Margaret's package
#For figures/plotting
library(ggplot2)
library(gridExtra)
library(ggsegDKT)
library(ggseg)
library(ggpubr)
library(glue)
library(ggcorrplot)
library(cowplot)
library(egg)
library(interactions)
library(growthstandards)
library(gtsummary)
library(svglite)
library(patchwork)


source(paste0(code_path, "load_data_fct.R"))
source("/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/longitudinal-models-abcd/long_centile_sim_plot.R")

#df to map variable names onto names displayed on figures
name_mapping <- data.frame(
  abbreviation = c(
     "smri_vol_cdk_totallh", "smri_vol_cdk_totalrh", "smri_vol_cdk_total",   
     "totalWM_cb", "totalWM_crb", "totalGM_crb",
     
    "smri_vol_scs_cbwmatterlh", "smri_vol_scs_ltventriclelh", "smri_vol_scs_inflatventlh",
    "smri_vol_scs_crbwmatterlh", "smri_vol_scs_crbcortexlh", "smri_vol_scs_tplh",
    "smri_vol_scs_caudatelh", "smri_vol_scs_putamenlh", "smri_vol_scs_pallidumlh",
    "smri_vol_scs_3rdventricle", "smri_vol_scs_4thventricle", "smri_vol_scs_bstem",
    "smri_vol_scs_hpuslh", "smri_vol_scs_amygdalalh", "smri_vol_scs_csf",
    "smri_vol_scs_lesionlh", "smri_vol_scs_aal", "smri_vol_scs_vedclh",
    
    "smri_vol_scs_cbwmatterrh", "smri_vol_scs_ltventriclerh", "smri_vol_scs_inflatventrh",
    "smri_vol_scs_crbwmatterrh", "smri_vol_scs_crbcortexrh", "smri_vol_scs_tprh",
    "smri_vol_scs_caudaterh", "smri_vol_scs_putamenrh", "smri_vol_scs_pallidumrh",
    "smri_vol_scs_hpusrh", "smri_vol_scs_amygdalarh", "smri_vol_scs_lesionrh",
    "smri_vol_scs_aar", "smri_vol_scs_vedcrh", 
    
    "smri_vol_scs_wmhint",
    "smri_vol_scs_wmhintlh", "smri_vol_scs_wmhintrh", 
    
    "smri_vol_scs_ccps",
    "smri_vol_scs_ccmidps", "smri_vol_scs_ccct", "smri_vol_scs_ccmidat",
    "smri_vol_scs_ccat", 
    
    "smri_vol_scs_wholeb", "smri_vol_scs_latventricles",
    "smri_vol_scs_allventricles", "smri_vol_scs_intracranialv", "smri_vol_scs_suprateialv",
    "smri_vol_scs_subcorticalgv",
    
    "smri_vol_cdk_banksstslh", "smri_vol_cdk_cdacatelh", "smri_vol_cdk_cdmdfrlh",
    "smri_vol_cdk_cuneuslh", "smri_vol_cdk_ehinallh", "smri_vol_cdk_fusiformlh",
    "smri_vol_cdk_ifpllh", "smri_vol_cdk_iftmlh", "smri_vol_cdk_ihcatelh",
    "smri_vol_cdk_locclh", "smri_vol_cdk_lobfrlh", "smri_vol_cdk_linguallh",
    "smri_vol_cdk_mobfrlh", "smri_vol_cdk_mdtmlh", "smri_vol_cdk_parahpallh",
    "smri_vol_cdk_paracnlh", "smri_vol_cdk_parsopclh", "smri_vol_cdk_parsobislh",
    "smri_vol_cdk_parstgrislh", "smri_vol_cdk_pericclh", "smri_vol_cdk_postcnlh",
    "smri_vol_cdk_ptcatelh", "smri_vol_cdk_precnlh", "smri_vol_cdk_pclh",
    "smri_vol_cdk_rracatelh", "smri_vol_cdk_rrmdfrlh", "smri_vol_cdk_sufrlh",
    "smri_vol_cdk_supllh", "smri_vol_cdk_sutmlh", "smri_vol_cdk_smlh",
    "smri_vol_cdk_frpolelh", "smri_vol_cdk_tmpolelh", "smri_vol_cdk_trvtmlh",
    "smri_vol_cdk_insulalh",
    
     "smri_vol_cdk_banksstsrh", "smri_vol_cdk_cdacaterh", "smri_vol_cdk_cdmdfrrh",
    "smri_vol_cdk_cuneusrh", "smri_vol_cdk_ehinalrh", "smri_vol_cdk_fusiformrh",
    "smri_vol_cdk_ifplrh", "smri_vol_cdk_iftmrh", "smri_vol_cdk_ihcaterh",
    "smri_vol_cdk_loccrh", "smri_vol_cdk_lobfrrh", "smri_vol_cdk_lingualrh",
    "smri_vol_cdk_mobfrrh", "smri_vol_cdk_mdtmrh", "smri_vol_cdk_parahpalrh",
    "smri_vol_cdk_paracnrh", "smri_vol_cdk_parsopcrh", "smri_vol_cdk_parsobisrh",
    "smri_vol_cdk_parstgrisrh", "smri_vol_cdk_periccrh", "smri_vol_cdk_postcnrh",
    "smri_vol_cdk_ptcaterh", "smri_vol_cdk_precnrh", "smri_vol_cdk_pcrh",
    "smri_vol_cdk_rracaterh", "smri_vol_cdk_rrmdfrrh", "smri_vol_cdk_sufrrh",
    "smri_vol_cdk_suplrh", "smri_vol_cdk_sutmrh", "smri_vol_cdk_smrh",
    "smri_vol_cdk_frpolerh", "smri_vol_cdk_tmpolerh", "smri_vol_cdk_trvtmrh",
    "smri_vol_cdk_insularh"
  ),
  region_name = c(
    "Left-Cerebral_Cortex", "Right-Cerebral-Cortex", "Total-Cerebral-Cortex",
    "Total_Cerebral-WM", "Total_Cerebellum-WM", "Total-Cerebellum-Cortex",
    
    "Left-Cerebral-White-Matter", "Left-Lateral-Ventricle", "Left-Inf-Lat-Vent",
    "Left-Cerebellum-White-Matter", "Left-Cerebellum-Cortex", "Left-Thalamus-Proper",
    "Left-Caudate", "Left-Putamen", "Left-Pallidum",
    "x3rd-ventricle", "x4th-ventricle", "Brain-Stem",
    "Left-Hippocampus", "Left-Amygdala", "Cerebrospinal-Fluid",
    "Left-Lesion", "Left-Accumbens-Area", "Left-VentralDC",
    
   "Right-Cerebral-White-Matter", "Right-Lateral-Ventricle", "Right-Inf-Lat-Vent",
    "Right-Cerebellum-White-Matter", "Right-Cerebellum-Cortex", "Right-Thalamus-Proper",
    "Right-Caudate", "Right-Putamen", "Right-Pallidum",
    "Right-Hippocampus", "Right-Amygdala",
    "Right-Lesion", "Right-Accumbens-Area", "Right-VentralDC",
    
    "WM-Hypointensities","Left-WM-Hypointensities", "Right-WM-Hypointensities",
   
   "CC_Posterior","CC_Mid_Posterior", "CC_Central", "CC_Mid_Anterior",
    "CC_Anterior", 
   
   "WholeBrain", "LatVentricles","AllVentricles", "IntracranialVolume", 
   "SupratentorialVolume", "SubcorticalGrayVolume",
   
     "lh_bankssts", "lh_caudalanteriorcingulate", "lh_caudalmiddlefrontal",
    "lh_cuneus", "lh_entorhinal", "lh_fusiform",
    "lh_inferiorparietal", "lh_inferiortemporal", "lh_isthmuscingulate",
    "lh_lateraloccipital", "lh_lateralorbitofrontal", "lh_lingual",
    "lh_medialorbitofrontal", "lh_middletemporal", "lh_parahippocampal",
    "lh_paracentral", "lh_parsopercularis", "lh_parsorbitalis",
    "lh_parstriangularis", "lh_pericalcarine", "lh_postcentral",
    "lh_posteriorcingulate", "lh_precentral", "lh_precuneus", "lh_rostralanteriorcingulate",
    "lh_rostralmiddlefrontal", "lh_superiorfrontal", "lh_superiorparietal",
    "lh_superiortemporal", "lh_supramarginal", "lh_frontalpole",
    "lh_temporalpole", "lh_transversetemporal", "lh_insula",
   
     "rh_bankssts", "rh_caudalanteriorcingulate", "rh_caudalmiddlefrontal",
    "rh_cuneus", "rh_entorhinal", "rh_fusiform",
    "rh_inferiorparietal", "rh_inferiortemporal", "rh_isthmuscingulate",
    "rh_lateraloccipital", "rh_lateralorbitofrontal", "rh_lingual",
    "rh_medialorbitofrontal", "rh_middletemporal", "rh_parahippocampal",
    "rh_paracentral", "rh_parsopercularis", "rh_parsorbitalis",
    "rh_parstriangularis", "rh_pericalcarine", "rh_postcentral",
    "rh_posteriorcingulate", "rh_precentral", "rh_precuneus", "rh_rostralanteriorcingulate",
    "rh_rostralmiddlefrontal", "rh_superiorfrontal", "rh_superiorparietal",
    "rh_superiortemporal", "rh_supramarginal", "rh_frontalpole",
    "rh_temporalpole", "rh_transversetemporal", "rh_insula"
    )
)

phenotypes_for_MC <- c(
    "smri_vol_scs_ltventriclelh", "smri_vol_scs_inflatventlh",
    "smri_vol_scs_crbwmatterlh", "smri_vol_scs_crbcortexlh", "smri_vol_scs_tplh",
    "smri_vol_scs_caudatelh", "smri_vol_scs_putamenlh", "smri_vol_scs_pallidumlh",
    "smri_vol_scs_3rdventricle", "smri_vol_scs_4thventricle", "smri_vol_scs_bstem",
    "smri_vol_scs_hpuslh", "smri_vol_scs_amygdalalh", "smri_vol_scs_aal", "smri_vol_scs_vedclh",
    
    "smri_vol_scs_ltventriclerh", "smri_vol_scs_inflatventrh",
    "smri_vol_scs_crbwmatterrh", "smri_vol_scs_crbcortexrh", "smri_vol_scs_tprh",
    "smri_vol_scs_caudaterh", "smri_vol_scs_putamenrh", "smri_vol_scs_pallidumrh",
    "smri_vol_scs_hpusrh", "smri_vol_scs_amygdalarh",
    "smri_vol_scs_aar", "smri_vol_scs_vedcrh", 
    
    "smri_vol_scs_ccps",
    "smri_vol_scs_ccmidps", "smri_vol_scs_ccct", "smri_vol_scs_ccmidat",
    "smri_vol_scs_ccat",
    
    "smri_vol_cdk_banksstslh", "smri_vol_cdk_cdacatelh", "smri_vol_cdk_cdmdfrlh",
    "smri_vol_cdk_cuneuslh", "smri_vol_cdk_ehinallh", "smri_vol_cdk_fusiformlh",
    "smri_vol_cdk_ifpllh", "smri_vol_cdk_iftmlh", "smri_vol_cdk_ihcatelh",
    "smri_vol_cdk_locclh", "smri_vol_cdk_lobfrlh", "smri_vol_cdk_linguallh",
    "smri_vol_cdk_mobfrlh", "smri_vol_cdk_mdtmlh", "smri_vol_cdk_parahpallh",
    "smri_vol_cdk_paracnlh", "smri_vol_cdk_parsopclh", "smri_vol_cdk_parsobislh",
    "smri_vol_cdk_parstgrislh", "smri_vol_cdk_pericclh", "smri_vol_cdk_postcnlh",
    "smri_vol_cdk_ptcatelh", "smri_vol_cdk_precnlh", "smri_vol_cdk_pclh",
    "smri_vol_cdk_rracatelh", "smri_vol_cdk_rrmdfrlh", "smri_vol_cdk_sufrlh",
    "smri_vol_cdk_supllh", "smri_vol_cdk_sutmlh", "smri_vol_cdk_smlh",
    "smri_vol_cdk_frpolelh", "smri_vol_cdk_tmpolelh", "smri_vol_cdk_trvtmlh",
    "smri_vol_cdk_insulalh",
    
     "smri_vol_cdk_banksstsrh", "smri_vol_cdk_cdacaterh", "smri_vol_cdk_cdmdfrrh",
    "smri_vol_cdk_cuneusrh", "smri_vol_cdk_ehinalrh", "smri_vol_cdk_fusiformrh",
    "smri_vol_cdk_ifplrh", "smri_vol_cdk_iftmrh", "smri_vol_cdk_ihcaterh",
    "smri_vol_cdk_loccrh", "smri_vol_cdk_lobfrrh", "smri_vol_cdk_lingualrh",
    "smri_vol_cdk_mobfrrh", "smri_vol_cdk_mdtmrh", "smri_vol_cdk_parahpalrh",
    "smri_vol_cdk_paracnrh", "smri_vol_cdk_parsopcrh", "smri_vol_cdk_parsobisrh",
    "smri_vol_cdk_parstgrisrh", "smri_vol_cdk_periccrh", "smri_vol_cdk_postcnrh",
    "smri_vol_cdk_ptcaterh", "smri_vol_cdk_precnrh", "smri_vol_cdk_pcrh",
    "smri_vol_cdk_rracaterh", "smri_vol_cdk_rrmdfrrh", "smri_vol_cdk_sufrrh",
    "smri_vol_cdk_suplrh", "smri_vol_cdk_sutmrh", "smri_vol_cdk_smrh",
    "smri_vol_cdk_frpolerh", "smri_vol_cdk_tmpolerh", "smri_vol_cdk_trvtmrh",
    "smri_vol_cdk_insularh"
  )

global_phenotypes <- c("smri_vol_cdk_total", "totalWM_cb", "smri_vol_scs_allventricles", "smri_vol_scs_subcorticalgv")

phenotypes_to_fit <- (c(phenotypes_for_MC, global_phenotypes, "smri_vol_scs_wholeb", "smri_vol_scs_wholeb", "totalWM_crb", "totalGM_crb", "smri_vol_scs_csf", "smri_vol_cdk_totallh", "smri_vol_cdk_totalrh", "smri_vol_scs_cbwmatterlh", "smri_vol_scs_cbwmatterrh"))

#save phenotypes not plotted with ggseg surface plots
plot_data_dkt <- merge(name_mapping, ggseg::dk, by.x = "region_name", by.y = "label")
plot_data_aseg <- merge(name_mapping, ggseg::aseg, by.x = "region_name", by.y = "label")

#Regions not matched in atlases that are tested
regions_notplotted <- name_mapping[!(name_mapping$region_name %in% plot_data_dkt$region_name) & !(name_mapping$region_name %in% plot_data_aseg$region_name),] %>% filter(abbreviation %in% phenotypes_for_MC)
write.csv(regions_notplotted, "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/regions_notplotted.csv")
```
#Create folders for saving output
```{r}
betas_folder <- paste0(save_path, "full_betas/")
if (!dir.exists(paste0(save_path, "full_betas/"))) dir.create(betas_folder)

dkt_folder <- paste0(save_path, "dkt_betas/")
if (!dir.exists(dkt_folder)) dir.create(dkt_folder)

tableOne_folder <- paste0(save_path, "table_one/")
if (!dir.exists(tableOne_folder)) dir.create(tableOne_folder)

supp_betas_table_folder <- paste0(save_path, "supplementary_betas_tables/")
if (!dir.exists(supp_betas_table_folder)) dir.create(supp_betas_table_folder)


figure_panels_folder <- paste0(save_path, "figure_panels/")
if (!dir.exists(figure_panels_folder)) dir.create(figure_panels_folder)

revisions_panels_folder <- paste0(save_path, "revisions_panels/")
if (!dir.exists(revisions_panels_folder)) dir.create(revisions_panels_folder)

extra_panels_folder <- paste0(save_path, "extra_panels/")
if (!dir.exists(extra_panels_folder)) dir.create(extra_panels_folder)

panel_save <- function(plot, filename, extension, folder){
  save_name <- paste0(folder, filename, ".", extension)
  ggsave(save_name, plot = plot, dpi = 300)
}
```
#Data Setup
Load all saved centiles data tables for t1, t2, and longitudinal
##T1
###Define files
```{r}
#file paths
splitOne_folder <- paste0(save_path, "gamlss_fits_t1_BCT_vol_splitONE_2.0/")
splitTwo_folder <- paste0(save_path,"gamlss_fits_t1_BCT_vol_splitTWO_2.0/")
sOne_fits_stats_filename <- paste0(splitOne_folder, "gamlss_fits_stats_all.csv")
sTwo_fits_stats_filename <- paste0(splitTwo_folder, "gamlss_fits_stats_all.csv")
splitOne_train_filename <- paste0(data_path, "abcd_long_Onetrain2025-09-26.csv")
splitTwo_train_filename <- paste0(data_path, "abcd_long_Twotrain2025-09-26.csv")
splitOne_test_filename <- paste0(data_path, "abcd_long_Onetest2025-09-26.csv")
splitTwo_test_filename <- paste0(data_path, "abcd_long_Twotest2025-09-26.csv")

t1.centiles_filename <- paste0(save_path, "centiles_calculated/t1.volume_centiles_2025-11-15.csv")
```
###Load data frames and clean them up
```{r}
#rename variables and select t1 scans
rename_vec <- c(age = "interview_age", sex = "sex_baseline", site = "site_id_l", nonSingleton = "nonSingleton")

t1.sOne_train_df <- read.csv(splitOne_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t1")
t1.sOne_train_df$sex <- as.factor(t1.sOne_train_df$sex)
t1.sOne_train_df$site <- as.factor(t1.sOne_train_df$site)
t1.sOne_train_df$age <- as.numeric(t1.sOne_train_df$age)

t1.sTwo_train_df <- read.csv(splitTwo_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t1")
t1.sTwo_train_df$sex <- as.factor(t1.sTwo_train_df$sex)
t1.sTwo_train_df$site <- as.factor(t1.sTwo_train_df$site)
t1.sTwo_train_df$age <- as.numeric(t1.sTwo_train_df$age)

t1.sOne_test_df <- read.csv(splitOne_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t1")
t1.sOne_test_df$sex <- as.factor(t1.sOne_test_df$sex)
t1.sOne_test_df$site <- as.factor(t1.sOne_test_df$site)
t1.sOne_test_df$age <- as.numeric(t1.sOne_test_df$age)

t1.sTwo_test_df <- read.csv(splitTwo_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t1")
t1.sTwo_test_df$sex <- as.factor(t1.sTwo_test_df$sex)
t1.sTwo_test_df$site <- as.factor(t1.sTwo_test_df$site)
t1.sTwo_test_df$age <- as.numeric(t1.sTwo_test_df$age)
```
###load saved centiles and merge with full test df (both sOne test and sTwo test)
```{r}
t1.centiles <- read.csv(t1.centiles_filename)  %>% select(-1)
names(t1.centiles) <- sub("_centiles", ".t1_centiles", names(t1.centiles))
t1.full_test_df <- rbind(t1.sOne_test_df,t1.sTwo_test_df)
t1.full_test_df <- merge(t1.centiles, t1.full_test_df, by = "src_subject_id")
```
###calculate birth weight percentile
```{r}
#convert birthweight into percentiles using the gitHub growthstandards package which use the intergrowth charts
t1.full_test_df <- t1.full_test_df %>%
  mutate(birth_weight_kg = birth_weight_oz_sum_adj * 0.0283495) %>%
  mutate(gestAge_days = gestAge*7+3) %>%
  mutate(sexLong = recode(sex, "M" = "Male", "F" = "Female"))

t1.full_test_df$bweight_percentile <- igb_wtkg2centile(t1.full_test_df$gestAge_days, t1.full_test_df$birth_weight_kg, sex = t1.full_test_df$sexLong)/100

t1.full_test_df$bw_category <- ifelse(t1.full_test_df$bweight_percentile < 0.10, "SGA", ifelse(t1.full_test_df$bweight_percentile > 0.90, "LGA", "AGA"))
t1.full_test_df$sexLong <- NULL
```
###QC for weight/height measures
```{r}
#keep original columns
t1.full_test_df$anthroweightcalc_og <- t1.full_test_df$anthroweightcalc
t1.full_test_df$anthroheightcalc_og <- t1.full_test_df$anthroheightcalc

#remove based on <30lbs, too low weight
#t1weight_mean <- mean(t1.full_test_df$anthroweightcalc, na.rm = TRUE)
#t1weight_sd <- sd(t1.full_test_df$anthroweightcalc, na.rm = TRUE)
t1.full_test_df$anthroweightcalc <- ifelse(t1.full_test_df$anthroweightcalc <30 , NA,
                                           t1.full_test_df$anthroweightcalc)

sum(is.na(t1.full_test_df$anthroweightcalc)) - sum(is.na(t1.full_test_df$anthroweightcalc_og)) #6 removed

#remove under 30 inches, two people at 0 and 4 inches removed, or over 80 inches, two removed
#t1height_mean <- mean(t1.full_test_df$anthroheightcalc, na.rm = TRUE)
#t1height_sd <- sd(t1.full_test_df$anthroheightcalc, na.rm = TRUE)
t1.full_test_df$anthroheightcalc <- ifelse(t1.full_test_df$anthroheightcalc < 30 | t1.full_test_df$anthroheightcalc > 80, NA,
                                           t1.full_test_df$anthroheightcalc)

sum(is.na(t1.full_test_df$anthroheightcalc)) - sum(is.na(t1.full_test_df$anthroheightcalc_og)) #4 removed

```
###encode full term as 38 weeks and 39 weeks
```{r}
t1.full_test_df$gestAge_38term <- ifelse(t1.full_test_df$gestAge == 40, 38, t1.full_test_df$gestAge)
t1.full_test_df$gestAge_39term <- ifelse(t1.full_test_df$gestAge == 40, 39, t1.full_test_df$gestAge)
```

### rename psychopathology subfactors
```{r}
t1.full_test_df <- t1.full_test_df %>%
  rename(psy = Factor1,
         mania = Factor2,
         conduct = Factor3,
         socanx = Factor4,
         att = Factor5,
         som = Factor6,
         overeat = Factor7,
         ptrauma = Factor8)
```
##T2
###Define files
```{r}
#file paths
t2.splitOne_folder <- paste0(save_path, "gamlss_fits_t2_BCT_vol_splitONE_2.0/")
t2.splitTwo_folder <- paste0(save_path, "gamlss_fits_t2_BCT_vol_splitTWO_2.0/")
t2.sOne_fits_stats_filename <- paste0(t2.splitOne_folder, "gamlss_fits_stats_all.csv")
t2.sTwo_fits_stats_filename <- paste0(t2.splitTwo_folder, "gamlss_fits_stats_all.csv")

splitOne_train_filename <-paste0(data_path, "abcd_long_Onetrain2025-09-26.csv")
splitTwo_train_filename <- paste0(data_path, "abcd_long_Twotrain2025-09-26.csv")
splitOne_test_filename <- paste0(data_path, "abcd_long_Onetest2025-09-26.csv")
splitTwo_test_filename <- paste0(data_path, "abcd_long_Twotest2025-09-26.csv")

t2.centiles_filename <- paste0(save_path, "centiles_calculated/t2.volume_centiles_2025-11-15.csv")
```
### Load data frames and clean them up
```{r}
#rename variables and select t2 scans
rename_vec <- c(age = "interview_age", sex = "sex_baseline", site = "site_id_l", nonSingleton = "nonSingleton")

t2.sOne_train_df <- read.csv(splitOne_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t2")
t2.sOne_train_df$sex <- as.factor(t2.sOne_train_df$sex)
t2.sOne_train_df$site <- as.factor(t2.sOne_train_df$site)
t2.sOne_train_df$age <- as.numeric(t2.sOne_train_df$age)

t2.sTwo_train_df <- read.csv(splitTwo_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t2")
t2.sTwo_train_df$sex <- as.factor(t2.sTwo_train_df$sex)
t2.sTwo_train_df$site <- as.factor(t2.sTwo_train_df$site)
t2.sTwo_train_df$age <- as.numeric(t2.sTwo_train_df$age)

t2.sOne_test_df <- read.csv(splitOne_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t2")
t2.sOne_test_df$sex <- as.factor(t2.sOne_test_df$sex)
t2.sOne_test_df$site <- as.factor(t2.sOne_test_df$site)
t2.sOne_test_df$age <- as.numeric(t2.sOne_test_df$age)

t2.sTwo_test_df <- read.csv(splitTwo_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec)) %>% 
  subset(eventname == "t2")
t2.sTwo_test_df$sex <- as.factor(t2.sTwo_test_df$sex)
t2.sTwo_test_df$site <- as.factor(t2.sTwo_test_df$site)
t2.sTwo_test_df$age <- as.numeric(t2.sTwo_test_df$age)
```
###load saved centiles and merge with full test df (both sOne test and sTwo test)
```{r}
t2.centiles <- read.csv(t2.centiles_filename)  %>% select(-1)
names(t2.centiles) <- sub("_centiles", ".t2_centiles", names(t2.centiles))
t2.full_test_df <- rbind(t2.sOne_test_df,t2.sTwo_test_df)
t2.full_test_df <- merge(t2.centiles, t2.full_test_df, by = "src_subject_id")
```

### Add variables for analyses
```{r}
# neonatal comp in this df is not calculated correctly, it is NA vector, take it from the t1 df
t2.full_test_df$neonatal_comp_total <- NULL

t2.full_test_df <- merge(t2.full_test_df, 
                           t1.full_test_df %>%
                             select(src_subject_id, gestAge_38term,bweight_percentile, neonatal_comp_total),
                           all.x = TRUE, all.y = FALSE, by = "src_subject_id")
```

###QC for weight/height measures
```{r}
#keep original columns
t2.full_test_df$anthroweightcalc_og <- t2.full_test_df$anthroweightcalc
t2.full_test_df$anthroheightcalc_og <- t2.full_test_df$anthroheightcalc

#remove based on >400lbs, one very high removed at 3580lbs
#t2weight_mean <- mean(t2.full_test_df$anthroweightcalc, na.rm = TRUE)
#t2weight_sd <- sd(t2.full_test_df$anthroweightcalc, na.rm = TRUE)
t2.full_test_df$anthroweightcalc <- ifelse(t2.full_test_df$anthroweightcalc > 400, NA,
                                           t2.full_test_df$anthroweightcalc)

sum(is.na(t2.full_test_df$anthroweightcalc)) - sum(is.na(t2.full_test_df$anthroweightcalc_og)) #1 removed

#remove under 30 inches, one person at 6.5 inches, over 80 inches - 3 people at 89+ inches removed
#t2height_mean <- mean(t2.full_test_df$anthroheightcalc, na.rm = TRUE)
#t2height_sd <- sd(t2.full_test_df$anthroheightcalc, na.rm = TRUE)
t2.full_test_df$anthroheightcalc <- ifelse(t2.full_test_df$anthroheightcalc < 30 | t2.full_test_df$anthroheightcalc > 80, NA,
                                           t2.full_test_df$anthroheightcalc)

sum(is.na(t2.full_test_df$anthroheightcalc)) - sum(is.na(t2.full_test_df$anthroheightcalc_og)) #4 removed

```

### rename psychopathology subfactors
```{r}
t2.full_test_df <- t2.full_test_df %>%
  rename(psy = Factor1,
         mania = Factor2,
         conduct = Factor3,
         socanx = Factor4,
         att = Factor5,
         som = Factor6,
         overeat = Factor7,
         ptrauma = Factor8)
```

##LONG
###Define files
```{r}
#file paths
long.splitOne_folder <- paste0(save_path, "gamlss_fits_long_vol_splitONE_2.0/")
long.splitTwo_folder <- paste0(save_path, "gamlss_fits_long_vol_splitTWO_2.0/")
long.sOne_fits_stats_filename <- paste0(long.splitOne_folder, "gamlss_fits_stats_all.csv")
long.sTwo_fits_stats_filename <- paste0(long.splitTwo_folder, "gamlss_fits_stats_all.csv")

long.splitOne_train_filename <- paste0(data_path, "abcd_wide_Onetrain2025-10-03.csv")
long.splitTwo_train_filename <- paste0(data_path, "abcd_wide_Twotrain2025-10-03.csv")
long.splitOne_test_filename <- paste0(data_path, "abcd_wide_Onetest2025-10-03.csv")
long.splitTwo_test_filename <- paste0(data_path, "abcd_wide_Twotest2025-10-03.csv")

long.centiles_filename <- paste0(save_path, "centiles_calculated/long.volume_centiles_2025-11-15.csv")
```
###Load data frames and clean them up
```{r}
#rename variables
rename_vec <- c(age = "interview_age.t2", sex = "sex_baseline", site = "site_id_l.t2")

long.sOne_train_df <- read.csv(long.splitOne_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
long.sOne_train_df$sex <- as.factor(long.sOne_train_df$sex)
long.sOne_train_df$site <- as.factor(long.sOne_train_df$site)
long.sOne_train_df$age <- as.numeric(long.sOne_train_df$age)

long.sTwo_train_df <- read.csv(long.splitTwo_train_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
long.sTwo_train_df$sex <- as.factor(long.sTwo_train_df$sex)
long.sTwo_train_df$site <- as.factor(long.sTwo_train_df$site)
long.sTwo_train_df$age <- as.numeric(long.sTwo_train_df$age)

long.sOne_test_df <- read.csv(long.splitOne_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
long.sOne_test_df$sex <- as.factor(long.sOne_test_df$sex)
long.sOne_test_df$site <- as.factor(long.sOne_test_df$site)
long.sOne_test_df$age <- as.numeric(long.sOne_test_df$age)

long.sTwo_test_df <- read.csv(long.splitTwo_test_filename)  %>% select(-1) %>% rename(., all_of(rename_vec))
long.sTwo_test_df$sex <- as.factor(long.sTwo_test_df$sex)
long.sTwo_test_df$site <- as.factor(long.sTwo_test_df$site)
long.sTwo_test_df$age <- as.numeric(long.sTwo_test_df$age)
```

###load saved centiles and merge with full test df (both sOne test and sTwo test)
```{r}
long.centiles <- read.csv(long.centiles_filename)  %>% select(-1)
names(long.centiles) <- sub(".t2_centiles", ".long_centiles", names(long.centiles))
long.full_test_df <- rbind(long.sOne_test_df,long.sTwo_test_df)
long.full_test_df <- merge(long.centiles, long.full_test_df, by = "src_subject_id")
```

### Add variables for analyses
```{r}
long.full_test_df <- merge(long.full_test_df, 
                           t1.full_test_df %>% rename (General_p.t1 = General_p,
                                            psy.t1 = psy, mania.t1 = mania, conduct.t1 = conduct,
                                            socanx.t1 = socanx, att.t1 = att, som.t1 = som,
                                            overeat.t1 = overeat, ptrauma.t1 = ptrauma,
                                            anthroheightcalc.t1 = anthroheightcalc, 
                                            anthroweightcalc.t1 = anthroweightcalc) %>% 
                             select(src_subject_id, birth_weight_oz_sum_adj, gestAge, gestAge_38term,
                                                      nonSingleton, bweight_percentile, neonatal_comp_total, 
                                                      bw_pgs, PC1, PC2, PC3, PC4, PC5, PC6, PC7, PC8, PC9, PC10,
                                                      General_p.t1, psy.t1, mania.t1, conduct.t1, socanx.t1, att.t1, 
                                                      som.t1, overeat.t1, ptrauma.t1, anthroweightcalc.t1, anthroheightcalc.t1),
                           all.x = TRUE, all.y = FALSE, by = "src_subject_id")

long.full_test_df <- merge(long.full_test_df, 
                           t2.full_test_df %>% rename (General_p.t2 = General_p,
                                            psy.t2 = psy, mania.t2 = mania, conduct.t2 = conduct,
                                            socanx.t2 = socanx, att.t2 = att, som.t2 = som,
                                            overeat.t2 = overeat, ptrauma.t2 = ptrauma,
                                            anthroheightcalc.t2 = anthroheightcalc, 
                                            anthroweightcalc.t2 = anthroweightcalc) %>% 
                             select(src_subject_id, General_p.t2, psy.t2, mania.t2, conduct.t2, socanx.t2, att.t2, 
                                                      som.t2, overeat.t2, ptrauma.t2, anthroweightcalc.t2, anthroheightcalc.t2),
                           all.x = TRUE, all.y = FALSE, by = "src_subject_id") %>%
    mutate(General_p_change = General_p.t2 - General_p.t1,
          psy_change = psy.t2 - psy.t1,  
          mania_change = mania.t2 - mania.t1,  
          conduct_change = conduct.t2 - conduct.t1,  
          socanx_change = socanx.t2 - socanx.t2, 
          att_change = att.t2 - att.t1,  
          som_change = som.t2 - som.t1,  
          overeat_change = overeat.t2 - overeat.t1,  
          ptrauma_change = ptrauma.t2 - ptrauma.t1)
```

## Add psychopathology change to t1 and t2 dfs for testing
```{r}
t1.full_test_df <- merge(t1.full_test_df, 
                         long.full_test_df %>% select(src_subject_id, grep("change", names(long.full_test_df))),
                         by = "src_subject_id", all.x = TRUE)
t2.full_test_df <- merge(t2.full_test_df, 
                         long.full_test_df %>% select(src_subject_id, grep("change", names(long.full_test_df))),
                         by = "src_subject_id", all.x = TRUE)
```

## Define vectors for names of centile vars for all models
```{r}
names <- names(t1.full_test_df)[grep(".t1_centiles", names(t1.full_test_df))] %>% sub(".t1_centiles","",.)
names.t1 <- paste0(names, ".t1_centiles")
names.t2 <- paste0(names, ".t2_centiles")
names.long <- paste0(names, ".long_centiles")
```

## Combine dataframes for centile-centile association analyses
To look at centile scores of the same subject across models.
```{r}
full.df <- merge(t1.full_test_df %>% rename(age.t1 = age, site.t1 = site, General_p.t1 = General_p) %>%
                   select(all_of(c("src_subject_id", "age.t1", "site.t1", "General_p.t1",
                                   names.t1, names, "split_ID", "sex", "gestAge", 
                                   "birth_weight_oz_sum_adj", "bw_pgs", "bweight_percentile", "neonatal_comp_total"))),
                  t2.full_test_df %>% rename(age.t2 = age, site.t2 = site, General_p.t2 = General_p) %>%
                   select(all_of(c("src_subject_id", "age.t2", "site.t2", "General_p.t2",
                                   names.t2, names))),
                 by = "src_subject_id", suffixes = c(".t1",".t2"))

#add longitudinal centiles
full.df <- merge(full.df, 
                 long.full_test_df %>% 
                   select(all_of(c("src_subject_id", names.long, "interscan_months_t1t2", "General_p_change"))),
                 by = "src_subject_id")
```

##SES Variables
###Load file
```{r}
abcd_demo_file <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/CSV/raw_tables_5.1/abcd_p_demo.csv"
abcd_demo <- read.csv(abcd_demo_file)
```

###Change encoding of variables
```{r}
abcd_demo <- abcd_demo %>%
  mutate(t1.income_label = case_when(
    demo_comb_income_v2 == 1 ~ "Less than $5,000",
    demo_comb_income_v2 == 2 ~ "$5,000 through $11,999",
    demo_comb_income_v2 == 3 ~ "$12,000 through $15,999",
    demo_comb_income_v2 == 4 ~ "$16,000 through $24,999",
    demo_comb_income_v2 == 5 ~ "$25,000 through $34,999",
    demo_comb_income_v2 == 6 ~ "$35,000 through $49,999",
    demo_comb_income_v2 == 7 ~ "$50,000 through $74,999",
    demo_comb_income_v2 == 8 ~ "$75,000 through $99,999",
    demo_comb_income_v2 == 9 ~ "$100,000 through $199,999",
    demo_comb_income_v2 == 10 ~ "$200,000 and greater",
    demo_comb_income_v2 == 999 ~ "Don't know",
    demo_comb_income_v2 == 777 ~ "Refuse to answer",
    TRUE ~ as.character(demo_comb_income_v2)  # Default case, if none of the above match
  ))

abcd_demo <- abcd_demo %>%
  mutate(t2.income_label = case_when(
    demo_comb_income_v2_l == 1 ~ "Less than $5,000",
    demo_comb_income_v2_l == 2 ~ "$5,000 through $11,999",
    demo_comb_income_v2_l == 3 ~ "$12,000 through $15,999",
    demo_comb_income_v2_l == 4 ~ "$16,000 through $24,999",
    demo_comb_income_v2_l == 5 ~ "$25,000 through $34,999",
    demo_comb_income_v2_l == 6 ~ "$35,000 through $49,999",
    demo_comb_income_v2_l == 7 ~ "$50,000 through $74,999",
    demo_comb_income_v2_l == 8 ~ "$75,000 through $99,999",
    demo_comb_income_v2_l == 9 ~ "$100,000 through $199,999",
    demo_comb_income_v2_l == 10 ~ "$200,000 and greater",
    demo_comb_income_v2_l == 999 ~ "Don't know",
    demo_comb_income_v2_l == 777 ~ "Refuse to answer",
    TRUE ~ as.character(demo_comb_income_v2_l)  # Default case, if none of the above match
  ))

abcd_demo <- abcd_demo %>%
  mutate(t1.prnt_edu_label = case_when(
    demo_prnt_ed_v2 %in% c(0:12) ~ "Didn't complete highschool",
    demo_prnt_ed_v2 %in% c(13,14) ~ "Completed highschool/GED",
    demo_prnt_ed_v2 %in% c(15:17) ~ "Some college/Associate's",
    demo_prnt_ed_v2 == 18 ~ "Bachelor's Degree",
    demo_prnt_ed_v2 %in% c(19:21) ~ "Advanced Degree",
    demo_prnt_ed_v2 == 999 ~ "Don't know",
    demo_prnt_ed_v2 == 777 ~ "Refuse to answer",
    TRUE ~ as.character(demo_prnt_ed_v2)  # Default case, if none of the above match
  ))

abcd_demo <- abcd_demo %>%
  mutate(t2.prnt_edu_label = case_when(
    demo_prnt_ed_v2_2yr_l %in% c(0:12) ~ "Didn't complete highschool",
    demo_prnt_ed_v2_2yr_l %in% c(13,14) ~ "Completed highschool/GED",
    demo_prnt_ed_v2_2yr_l %in% c(22,23,16,17) ~ "Some college/Associate's",
    demo_prnt_ed_v2_2yr_l == 18 ~ "Bachelor's Degree",
    demo_prnt_ed_v2_2yr_l %in% c(19:21) ~ "Advanced Degree",
    demo_prnt_ed_v2_2yr_l == 999 ~ "Don't know",
    demo_prnt_ed_v2_2yr_l == 777 ~ "Refuse to answer",
    TRUE ~ as.character(demo_prnt_ed_v2_2yr_l)  # Default case, if none of the above match
  ))
```

### Deal with NAs and Factor Ordering
set "no response" to NA
make the factors ordered
```{r}
abcd_demo$t1.income_label_NAs <- ifelse(abcd_demo$t1.income_label %in% c("Don't know", "Refuse to answer"), NA, abcd_demo$t1.income_label)
abcd_demo$t2.income_label_NAs <- ifelse(abcd_demo$t2.income_label %in% c("Don't know", "Refuse to answer"), NA, abcd_demo$t2.income_label)

abcd_demo$t1.prnt_edu_label_NAs <- ifelse(abcd_demo$t1.prnt_edu_label %in% c("Don't know", "Refuse to answer"), NA, abcd_demo$t1.prnt_edu_label)
abcd_demo$t2.prnt_edu_label_NAs <- ifelse(abcd_demo$t2.prnt_edu_label %in% c("Don't know", "Refuse to answer"), NA, abcd_demo$t2.prnt_edu_label)

t2.abcd_demo <- abcd_demo %>% filter(eventname == "2_year_follow_up_y_arm_1") %>%
  filter(src_subject_id %in% full.df$src_subject_id)
t1.abcd_demo <- abcd_demo %>% filter(eventname == "baseline_year_1_arm_1")%>%
  filter(src_subject_id %in% full.df$src_subject_id)

income_levels_order <- c("Less than $5,000","$5,000 through $11,999","$12,000 through $15,999", 
  "$16,000 through $24,999","$25,000 through $34,999","$35,000 through $49,999",
  "$50,000 through $74,999","$75,000 through $99,999","$100,000 through $199,999",
  "$200,000 and greater")
prnt_edu_levels_order <- c("Didn't complete highschool", "Completed highschool/GED", "Some college/Associate's", "Bachelor's Degree", "Advanced Degree")
```

### Add variables to data frames
```{r}
#Add Income Levels to full df and to t1 / t2 dfs. Set level orders for ordered factor
##T1 Income
t1.full_test_df <- merge(t1.full_test_df, t1.abcd_demo %>% select(c("src_subject_id", "t1.income_label_NAs", "t1.income_label")) %>% rename(t1.income_ordered = t1.income_label_NAs, t1.income_label = t1.income_label), by = "src_subject_id", all.x = TRUE)
t1.full_test_df$t1.income_ordered <- factor(t1.full_test_df$t1.income_ordered, levels = income_levels_order, ordered = TRUE)

t2.full_test_df <- merge(t2.full_test_df, t1.abcd_demo %>% select(c("src_subject_id", "t1.income_label_NAs", "t1.income_label")) %>% rename(t1.income_ordered = t1.income_label_NAs), by = "src_subject_id", all.x = TRUE)
t2.full_test_df$t1.income_ordered <- factor(t2.full_test_df$t1.income_ordered, levels = income_levels_order, ordered = TRUE)

long.full_test_df <- merge(long.full_test_df, t1.abcd_demo %>% select(c("src_subject_id", "t1.income_label_NAs", "t1.income_label")) %>% rename(t1.income_ordered = t1.income_label_NAs), by = "src_subject_id", all.x = TRUE)
long.full_test_df$t1.income_ordered <- factor(long.full_test_df$t1.income_ordered, levels = income_levels_order, ordered = TRUE)

##T2 Income
t1.full_test_df <- merge(t1.full_test_df, t2.abcd_demo %>% select(c("src_subject_id", "t2.income_label_NAs", "t2.income_label")) %>% rename(t2.income_ordered = t2.income_label_NAs), by = "src_subject_id", all.x = TRUE)
t1.full_test_df$t2.income_ordered <- factor(t1.full_test_df$t2.income_ordered, levels = income_levels_order, ordered = TRUE)

t2.full_test_df <- merge(t2.full_test_df, t2.abcd_demo %>% select(c("src_subject_id", "t2.income_label_NAs", "t2.income_label")) %>% rename(t2.income_ordered = t2.income_label_NAs, t2.income_label = t2.income_label), by = "src_subject_id", all.x = TRUE)
t2.full_test_df$t2.income_ordered <- factor(t2.full_test_df$t2.income_ordered, levels = income_levels_order, ordered = TRUE)

long.full_test_df <- merge(long.full_test_df, t2.abcd_demo %>% select(c("src_subject_id", "t2.income_label_NAs", "t2.income_label")) %>% rename(t2.income_ordered = t2.income_label_NAs), by = "src_subject_id", all.x = TRUE)
long.full_test_df$t2.income_ordered <- factor(long.full_test_df$t2.income_ordered, levels = income_levels_order, ordered = TRUE)

#Add Parent Education Levels to full df and to t1 / t2 dfs. Set level orders for ordered factor
##T1 Parent Education
t1.full_test_df <- merge(t1.full_test_df, t1.abcd_demo %>% select(c("src_subject_id", "t1.prnt_edu_label_NAs", "t1.prnt_edu_label")) %>% rename(t1.prnt_edu_ordered = t1.prnt_edu_label_NAs, t1.prnt_edu_label = t1.prnt_edu_label), by = "src_subject_id", all.x = TRUE)
t1.full_test_df$t1.prnt_edu_ordered <- factor(t1.full_test_df$t1.prnt_edu_ordered, levels = prnt_edu_levels_order, ordered = TRUE)

t2.full_test_df <- merge(t2.full_test_df, t1.abcd_demo %>% select(c("src_subject_id", "t1.prnt_edu_label_NAs", "t1.prnt_edu_label")) %>% rename(t1.prnt_edu_ordered = t1.prnt_edu_label_NAs), by = "src_subject_id", all.x = TRUE)
t2.full_test_df$t1.prnt_edu_ordered <- factor(t2.full_test_df$t1.prnt_edu_ordered, levels = prnt_edu_levels_order, ordered = TRUE)

long.full_test_df <- merge(long.full_test_df, t1.abcd_demo %>% select(c("src_subject_id", "t1.prnt_edu_label_NAs", "t1.prnt_edu_label")) %>% rename(t1.prnt_edu_ordered = t1.prnt_edu_label_NAs), by = "src_subject_id", all.x = TRUE)
long.full_test_df$t1.prnt_edu_ordered <- factor(long.full_test_df$t1.prnt_edu_ordered, levels = prnt_edu_levels_order, ordered = TRUE)

##T2 Parent Education
t1.full_test_df <- merge(t1.full_test_df, t2.abcd_demo %>% select(c("src_subject_id", "t2.prnt_edu_label_NAs", "t2.prnt_edu_label")) %>% rename(t2.prnt_edu_ordered = t2.prnt_edu_label_NAs), by = "src_subject_id", all.x = TRUE)
t1.full_test_df$t2.prnt_edu_ordered <- factor(t1.full_test_df$t2.prnt_edu_ordered, levels = prnt_edu_levels_order, ordered = TRUE)

t2.full_test_df <- merge(t2.full_test_df, t2.abcd_demo %>% select(c("src_subject_id", "t2.prnt_edu_label_NAs", "t2.prnt_edu_label")) %>% rename(t2.prnt_edu_ordered = t2.prnt_edu_label_NAs, t2.prnt_edu_label = t2.prnt_edu_label), by = "src_subject_id", all.x = TRUE)
t2.full_test_df$t2.prnt_edu_ordered <- factor(t2.full_test_df$t2.prnt_edu_ordered, levels = prnt_edu_levels_order, ordered = TRUE)

long.full_test_df <- merge(long.full_test_df, t2.abcd_demo %>% select(c("src_subject_id", "t2.prnt_edu_label_NAs", "t2.prnt_edu_label")) %>% rename(t2.prnt_edu_ordered = t2.prnt_edu_label_NAs), by = "src_subject_id", all.x = TRUE)
long.full_test_df$t2.prnt_edu_ordered <- factor(long.full_test_df$t2.prnt_edu_ordered, levels = prnt_edu_levels_order, ordered = TRUE)
```

### Create a combo variable for income and parental education, using t1 if available and t2 if t1 is not available
```{r}
#T1
t1.full_test_df$combo.income_ordered <- t1.full_test_df$t1.income_ordered
t1.full_test_df$combo.income_ordered[is.na(t1.full_test_df$t1.income_ordered)] <- t1.full_test_df$t2.income_ordered[is.na(t1.full_test_df$t1.income_ordered)]

t1.full_test_df$combo.prnt_edu_ordered <- t1.full_test_df$t1.prnt_edu_ordered
t1.full_test_df$combo.prnt_edu_ordered[is.na(t1.full_test_df$t1.prnt_edu_ordered)] <- t1.full_test_df$t2.prnt_edu_ordered[is.na(t1.full_test_df$t1.prnt_edu_ordered)]

#T2
t2.full_test_df$combo.income_ordered <- t2.full_test_df$t1.income_ordered
t2.full_test_df$combo.income_ordered[is.na(t2.full_test_df$t1.income_ordered)] <- t2.full_test_df$t2.income_ordered[is.na(t2.full_test_df$t1.income_ordered)]

t2.full_test_df$combo.prnt_edu_ordered <- t2.full_test_df$t1.prnt_edu_ordered
t2.full_test_df$combo.prnt_edu_ordered[is.na(t2.full_test_df$t1.prnt_edu_ordered)] <- t2.full_test_df$t2.prnt_edu_ordered[is.na(t2.full_test_df$t1.prnt_edu_ordered)]

#LONG
long.full_test_df$combo.income_ordered <- long.full_test_df$t1.income_ordered
long.full_test_df$combo.income_ordered[is.na(long.full_test_df$t1.income_ordered)] <- long.full_test_df$t2.income_ordered[is.na(long.full_test_df$t1.income_ordered)]

long.full_test_df$combo.prnt_edu_ordered <- long.full_test_df$t1.prnt_edu_ordered
long.full_test_df$combo.prnt_edu_ordered[is.na(long.full_test_df$t1.prnt_edu_ordered)] <- long.full_test_df$t2.prnt_edu_ordered[is.na(long.full_test_df$t1.prnt_edu_ordered)]
```

# Make Table One and Supplementary Table
by timepoint and split-half
```{r}
t1.full_test_df$table_group <- paste0(t1.full_test_df$matched_group, "_t1")
t2.full_test_df$table_group <- paste0(t2.full_test_df$matched_group, "_t2")

table_vars <- c("table_group", "age", "sex", "gestAge", "bw_pgs", "birth_weight_oz_sum_adj",
"neonatal_comp_total", "General_p", "psy", "mania", "conduct", "socanx", "att", "som", "overeat", "ptrauma",
"anthroweightcalc", "anthroheightcalc", "combo.income_ordered", "combo.prnt_edu_ordered")

table1_df <- rbind(t1.full_test_df %>% select(all_of(table_vars)), 
                   t2.full_test_df %>% select(all_of(table_vars))) %>%
  mutate(age_years = age/12) %>%
  mutate(bwpgs_zscore = as.numeric(scale(bw_pgs))) %>%
  mutate(neonatal_comp_total = as.numeric(neonatal_comp_total)) %>%
  mutate(table_group = factor(table_group, levels = c("1_t1", "2_t1", "1_t2", "2_t2"))) %>%
  arrange(table_group)

# Create Table 1 with additional variables
table1 <- table1_df %>%
  select(table_group, age_years, sex, gestAge, bwpgs_zscore, birth_weight_oz_sum_adj, neonatal_comp_total, General_p) %>%
  tbl_summary(
    by = table_group,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = c(bwpgs_zscore, General_p) ~ function(x) format(x, digits = 10),
    missing = "always",
    missing_stat = "{N_nonmiss}", 
    missing_text = "N known"
  ) %>%
  bold_labels()

# View Table
table1

table1_save <- as_tibble(table1)

write.csv(table1_save, paste0(tableOne_folder, "Table_One_", Sys.Date(),".csv"))

#Neonatal comp total is added as a categorical (I can't fix the function above to see it as continuous), want to include as a continuous -- calculate mean and SD manually here.
table1_df %>%
  select(c("table_group", "neonatal_comp_total")) %>%
  group_by(table_group) %>%
 summarise(mean_value = mean(neonatal_comp_total, na.rm = TRUE),
    sd_value = sd(neonatal_comp_total, na.rm = TRUE))

# Create Supplemental Table 1 with additional variables
table1_sup <- table1_df %>%
  select(table_group, anthroweightcalc, anthroheightcalc, combo.income_ordered, combo.prnt_edu_ordered, bw_pgs, neonatal_comp_total) %>%
  tbl_summary(
    by = table_group,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = c(bw_pgs) ~ function(x) format(x, digits = 3, scientific = TRUE),
    missing = "always",
    missing_stat = "{N_nonmiss}", 
    missing_text = "N known"
  ) %>%
  bold_labels()

# View Table
table1_sup

table1_sup_save <- as_tibble(table1_sup)

write.csv(table1_sup_save, paste0(tableOne_folder, "Table_One_Supplement_", Sys.Date(), ".csv"))
```

# Make Supplemental Table with psychopathology sub-factors
```{r}
table1_sup_psyfac <- table1_df %>%
  select(table_group, psy, mania, conduct, socanx, att, som, overeat, ptrauma) %>%
  tbl_summary(
    by = table_group,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "always",
    missing_stat = "{N_nonmiss}", 
    missing_text = "N known"
  ) %>%
  bold_labels()

# View Table
table1_sup_psyfac

table1_sup_psyfac_save <- as_tibble(table1_sup_psyfac)

write.csv(table1_sup_psyfac_save, paste0(tableOne_folder, "Table_One_Supplement_PsyFac", Sys.Date(), ".csv"))
```
# Prep full.df with zscores
## Set 0/1 centiles to min/max value to avoid Inf/-Inf
numbers from gamlssTools pkg
```{r}
qnorm_vars <- c(names.t1, names.t2, names.long, "bweight_percentile")

for (x in qnorm_vars){
  full.df[full.df[,x] == 0 & !is.na(full.df[,x]), x] <- 0.0000000000000000000000001
  full.df[full.df[,x] == 1 & !is.na(full.df[,x]), x] <- 0.99999999999999994
}
```
 
## zscore centiles (new df)
```{r, echo = FALSE}
# zscore centiles and other variables of interest
df.zscore <- full.df %>%
  mutate(across(all_of(qnorm_vars), qnorm)) %>%
  mutate(across(all_of(c("gestAge", "birth_weight_oz_sum_adj", "bw_pgs", "bweight_percentile", "neonatal_comp_total")), 
                ~as.numeric(scale(.x))))

# set sex as factor
df.zscore$sex <- as.factor(df.zscore$sex)
```

# Load Global Phenotype Models
##t1 global phenotype models
```{r}
#load saved fits
t1.all_fits_filename <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/t1_BCT_fits.csv"
t1.all_fits <- read.csv(t1.all_fits_filename) %>% select(-1) %>% filter(phenotype %in% c(global_phenotypes, "smri_vol_scs_wholeb"))
t1.sOne_fits <- t1.all_fits %>% filter(split_no == 1)
t1.sTwo_fits <- t1.all_fits %>% filter(split_no == 2)

#Choose best fits based on lowest BIC
t1.sOne_best <- t1.sOne_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
 slice_min(BIC, n = 1)

t1.sTwo_best <- t1.sTwo_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
  slice_min(BIC, n = 1)

#Load models for the best model for each phenotype
t1.sOne_best_list <- paste0(splitOne_folder,"gamlssFIT_",t1.sOne_best$model_index,".rds")
t1.best_models_One <- lapply(t1.sOne_best_list, readRDS)
t1.sTwo_best_list <- paste0(splitTwo_folder,"gamlssFIT_",t1.sTwo_best$model_index,".rds")
t1.best_models_Two <- lapply(t1.sTwo_best_list, readRDS)

names(t1.best_models_One) <- t1.sOne_best$phenotype
names(t1.best_models_Two) <- t1.sTwo_best$phenotype
```
##t2 global phenotype models
```{r}
#load saved fits
t2.all_fits_filename <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/t2_BCT_fits.csv"
t2.all_fits <- read.csv(t2.all_fits_filename) %>% select(-1) %>% filter(phenotype %in% c(global_phenotypes, "smri_vol_scs_wholeb"))
t2.sOne_fits <- t2.all_fits %>% filter(split_no == 1)
t2.sTwo_fits <- t2.all_fits %>% filter(split_no == 2)

#Choose best fits based on lowest BIC
t2.sOne_best <- t2.sOne_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
 slice_min(BIC, n = 1)

t2.sTwo_best <- t2.sTwo_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
  slice_min(BIC, n = 1)

#Load models for the best model for each phenotype
t2.sOne_best_list <- paste0(t2.splitOne_folder,"gamlssFIT_",t2.sOne_best$model_index,".rds")
t2.best_models_One <- lapply(t2.sOne_best_list, readRDS)
t2.sTwo_best_list <- paste0(t2.splitTwo_folder,"gamlssFIT_",t2.sTwo_best$model_index,".rds")
t2.best_models_Two <- lapply(t2.sTwo_best_list, readRDS)

names(t2.best_models_One) <- t2.sOne_best$phenotype
names(t2.best_models_Two) <- t2.sTwo_best$phenotype
```
##longitudinal global phenotype models
```{r}
global_t2 <- paste0(c(global_phenotypes, "smri_vol_scs_wholeb"), ".t2")
#load saved fits
long.all_fits_filename <- "/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/2.0_results/long_fits.csv"
long.all_fits <- read.csv(long.all_fits_filename) %>% select(-1) %>% filter(phenotype %in% global_t2)
long.sOne_fits <- long.all_fits %>% filter(split_no == 1)
long.sTwo_fits <- long.all_fits %>% filter(split_no == 2)

#Choose best fits based on lowest BIC
long.sOne_best <- long.sOne_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
 slice_min(BIC, n = 1)

long.sTwo_best <- long.sTwo_fits %>%
  filter(convergence_warn_end == 0) %>%
  group_by(phenotype) %>%
  slice_min(BIC, n = 1)

#Load models for the best model for each phenotype
long.sOne_best_list <- paste0(long.splitOne_folder,"gamlssFIT_",long.sOne_best$model_index,".rds")
long.best_models_One <- lapply(long.sOne_best_list, readRDS)
long.sTwo_best_list <- paste0(long.splitTwo_folder,"gamlssFIT_",long.sTwo_best$model_index,".rds")
long.best_models_Two <- lapply(long.sTwo_best_list, readRDS)

names(long.best_models_One) <- long.sOne_best$phenotype
names(long.best_models_Two) <- long.sTwo_best$phenotype
```
<!-- # Prepare DFs for Analysis -->

<!-- ## Set 0/1 centiles to min/max value to avoid Inf/-Inf -->
<!-- numbers from gamlssTools pkg -->
<!-- include weight/height centiles here later (when properly calculated and exported from arcus) -->
<!-- ```{r} -->
<!-- qnorm_vars <- c(names.t1, names.t2, names.long, "bweight_percentile") -->

<!-- for (x in qnorm_vars){ -->
<!--   full.df[full.df[,x] == 0 & !is.na(full.df[,x]), x] <- 0.0000000000000000000000001 -->
<!--   full.df[full.df[,x] == 1 & !is.na(full.df[,x]), x] <- 0.99999999999999994 -->
<!-- } -->
<!-- ``` -->

<!-- ## zscore centiles (new df) - ->
<!-- ```{r, echo = FALSE} -->
<!-- # zscore centiles and other variables of interest -->
<!-- df.zscore <- full.df %>% -->
<!--   mutate(across(all_of(qnorm_vars), qnorm)) %>% -->
<!--   mutate(across(all_of(c("gestAge", "gestAge_38term", "gestAge_39term", "birth_weight_oz_sum_adj",  -->
<!--                          "neonatal_comp_total", "bw_pgs", "ga_pgs",  -->
<!--                          "PC1", "PC2", "PC3", "PC4", "PC5","PC6","PC6",  -->
<!--                          "PC7", "PC7", "PC8", "PC9", "PC10", -->
<!--                          "parental_education", "General_p.t1", "General_p.t2", "General_p_change", -->
<!--                          "psy.t1", "psy.t2", "psy_change", "mania.t1", "mania.t2", "mania_change", -->
<!--                          "conduct.t1", "conduct.t2", "conduct_change", "socanx.t1", "socanx.t2", -->
<!--                          "socanx_change", "att.t1", "att.t2", "att_change", "som.t1", "som.t2", -->
<!--                          "som_change", "overeat.t1", "overeat.t2", "overeat_change", "ptrauma.t1", -->
<!--                          "ptrauma.t2", "ptrauma_change")),  -->
<!--                 ~as.numeric(scale(.x)))) -->

<!-- # set sex and preterm as factor -->
<!-- df.zscore$sex <- as.factor(df.zscore$sex) -->
<!-- df.zscore$preterm <- as.factor(df.zscore$preterm) -->

<!-- # # zscore centiles and other variables of interest -->
<!-- # df.zscore.t1 <- t1.full_test_df %>% -->
<!-- #   mutate(across(all_of(c(names.t1, "bweight_percentile")), qnorm)) %>% -->
<!-- #   mutate(across(all_of(c("gestAge", "birth_weight_oz_sum_adj",  -->
<!-- #                          "neonatal_comp_total", "bw_pgs", "ga_pgs",  -->
<!-- #                          "PC1", "PC2", "PC3", "PC4", "PC5","PC6","PC6",  -->
<!-- #                          "PC7", "PC7", "PC8", "PC9", "PC10", -->
<!-- #                          "parental_education", "General_p.t1", "General_p.t2", "General_p_change", -->
<!-- #                          "psy.t1", "psy.t2", "psy_change", "mania.t1", "mania.t2", "mania_change", -->
<!-- #                          "conduct.t1", "conduct.t2", "conduct_change", "socanx.t1", "socanx.t2", -->
<!-- #                          "socanx_change", "att.t1", "att.t2", "att_change", "som.t1", "som.t2", -->
<!-- #                          "som_change", "overeat.t1", "overeat.t2", "overeat_change", "ptrauma.t1", -->
<!-- #                          "ptrauma.t2", "ptrauma_change")),  -->
<!-- #                 ~as.numeric(scale(.x)))) -->
<!-- #  -->
<!-- # # set sex and preterm as factor -->
<!-- # df.zscore.t1$sex <- as.factor(df.zscore.t1$sex) -->
<!-- # df.zscore.t1$preterm <- as.factor(df.zscore.t1$preterm) -->

<!-- # # zscore centiles and other variables of interest -->
<!-- # df.zscore.long <- long.full_test_df %>% -->
<!-- #   mutate(across(all_of(c(names.long, "bweight_percentile")), qnorm)) %>% -->
<!-- #   mutate(across(all_of(c("gestAge", "birth_weight_oz_sum_adj",  -->
<!-- #                          "neonatal_comp_total", "bw_pgs", -->
<!-- #                          "PC1", "PC2", "PC3", "PC4", "PC5","PC6","PC6",  -->
<!-- #                          "PC7", "PC7", "PC8", "PC9", "PC10", -->
<!-- #                          "General_p.t1", "General_p.t2", "General_p_change", -->
<!-- #                          "psy.t1", "psy.t2", "psy_change", "mania.t1", "mania.t2", "mania_change", -->
<!-- #                          "conduct.t1", "conduct.t2", "conduct_change", "socanx.t1", "socanx.t2", -->
<!-- #                          "socanx_change", "att.t1", "att.t2", "att_change", "som.t1", "som.t2", -->
<!-- #                          "som_change", "overeat.t1", "overeat.t2", "overeat_change", "ptrauma.t1", -->
<!-- #                          "ptrauma.t2", "ptrauma_change")),  -->
<!-- #                 ~as.numeric(scale(.x)))) -->
<!-- #  -->
<!-- # # set sex and preterm as factor -->
<!-- # df.zscore.long$sex <- as.factor(df.zscore.long$sex) -->
<!-- # df.zscore.long$preterm <- as.factor(df.zscore.long$preterm) -->
<!-- ``` -->

<!-- ## find Inf cells, sanity check, should be none -->
<!-- ```{r, echo = FALSE} -->
<!-- inf_nan_table <- as.data.frame(t(sapply(df.zscore[c(qnorm_vars)], function(x){ -->
<!--   c(inf = sum(is.infinite(x)), nan = sum(is.nan(x))) -->
<!-- }))) -->
<!-- max(inf_nan_table$inf) -->
<!-- max(inf_nan_table$nan) -->

<!-- inf_nan_table %>% filter(inf > 0) -->
<!-- ``` -->

<!-- ## Store non-normalized distribution stats to convert back for plots -->
<!-- ```{r} -->
<!-- stats <- full.df %>% -->
<!--   dplyr::summarise(across(all_of(c("gestAge", "birth_weight_oz_sum_adj",  -->
<!--                          "neonatal_comp_total", "bw_pgs", "ga_pgs",  -->
<!--                          "PC1", "PC2", "PC3", "PC4", "PC5","PC6","PC6",  -->
<!--                          "PC7", "PC7", "PC8", "PC9", "PC10", -->
<!--                          "parental_education", "General_p.t1", "General_p.t2", "General_p_change", -->
<!--                          "psy.t1", "psy.t2", "psy_change", "mania.t1", "mania.t2", "mania_change", -->
<!--                          "conduct.t1", "conduct.t2", "conduct_change", "socanx.t1", "socanx.t2", -->
<!--                          "socanx_change", "att.t1", "att.t2", "att_change", "som.t1", "som.t2", -->
<!--                          "som_change", "overeat.t1", "overeat.t2", "overeat_change", "ptrauma.t1", -->
<!--                          "ptrauma.t2", "ptrauma_change")), -->
<!--                          list(mean = ~ mean(.x, na.rm = TRUE), -->
<!--                               sd = ~ sd(.x, na.rm = TRUE)), -->
<!--                          .names = "{.col}__{.fn}")) %>% -->
<!--   pivot_longer(everything(), -->
<!--                names_to = c("var", ".value"), -->
<!--                names_pattern = "^(.*)__(mean|sd)$") -->

<!-- mean_lookup <- setNames(stats$mean, stats$var) -->
<!-- sd_lookup <- setNames(stats$sd, stats$var) -->
<!-- ``` -->

# BW - PGSBW corr
```{r}
cor.test(full.df$birth_weight_oz_sum_adj, full.df$bw_pgs)
```
# Birth Factors ~ P-factor corr

## General P factor
```{r}
cor_mat <- cor(full.df[,c("gestAge", "birth_weight_oz_sum_adj", "bw_pgs", "neonatal_comp_total", "General_p.t1", "General_p.t2", "General_p_change")], use = "pairwise.complete.obs")

p_mat <- cor_pmat(full.df[,c("gestAge", "birth_weight_oz_sum_adj", "bw_pgs", "neonatal_comp_total", "General_p.t1", "General_p.t2", "General_p_change")], use = "pairwise.complete.obs")

(bf_gp_corr_plot <- ggcorrplot(
  cor_mat,
  p.mat = p_mat,
  insig = "blank",
  lab = TRUE
))

if(save_figures == TRUE){
panel_save(bf_gp_corr_plot, "bf-gp-cor-matrix", "svg", extra_panels_folder)
}
```

#Centiles Across Models
## Plot centiles against each other & calculate correlations for each phenotype
```{r}
length_names <- length(grep(".t1_centiles", names(full.df)))
#Ensure all names are in the same order
names <- names(full.df)[grep(".t1_centiles", names(full.df))] %>% sub(".t1_centiles","",.)
names.t1 <- paste0(names, ".t1_centiles")
names.t2 <- paste0(names, ".t2_centiles")
names.long <- paste0(names, ".long_centiles")
plots <- vector(mode = "list", length = length_names)
plots.long <- vector(mode = "list", length = length_names)
plots.t2 <- vector(mode = "list", length = length_names)
plots.t1 <- vector(mode = "list", length = length_names)
plots.t1_diff <- vector(mode = "list", length = length_names)
df <- data.frame(matrix(ncol = 4, nrow = dim(full.df)[1]))
colnames(df) <- c("t1.vector", "t2.vector", "diff.vector", "long.vector")
cor_df <- data.frame(matrix(ncol = 13, nrow = length_names))
colnames(cor_df) <- c("t1_t2.cor", "t1_t2.p", "t1_t2.p_adj", "t1_long.cor", "t1_long.p", "t1_long.p_adj","t2_long.cor", "t2_long.p", "t2_long.p_adj", "t1_diff.cor", "t1_diff.p", "t1_diff.p_adj",
                      "label")
for(i in 1:length_names){
  df$t1.vector.qnorm <- df.zscore[,names.t1[i]]
  df$t1.vector <- full.df[,names.t1[i]]
  df$t2.vector.qnorm <- df.zscore[,names.t2[i]]
  df$t2.vector <- full.df[,names.t2[i]]
  col_name <- paste0(names[i], ".diff")
  full.df[[col_name]] <- df$t2.vector-df$t1.vector
  col_name_qnorm <- paste0(names[i], ".diff.qnorm")
  df.zscore[[col_name_qnorm]] <- df$t2.vector.qnorm-df$t1.vector.qnorm
  df$diff.vector <- df$t2.vector-df$t1.vector
  df$diff.vector.qnorm <- df$t2.vector.qnorm-df$t1.vector.qnorm
  df$long.vector.qnorm <- df.zscore[,names.long[i]]
  df$long.vector <- full.df[,names.long[i]]
  title <-  as.character(name_mapping$region_name[match(names[i],  name_mapping$abbreviation)])
  plots[[i]] <- ggplot(data = df, mapping = aes(x = t1.vector, y = t2.vector, color = long.vector)) + geom_point(aes(alpha = 0.5)) + ggtitle(title)
  names(plots)[i] <- title
  plots.long[[i]]<- ggplot(data = df, mapping = aes(x = diff.vector, y = long.vector, color = t1.vector)) + geom_point(aes(alpha = 0.3)) + ggtitle(title) + xlab("t2-t1 centiles")
  names(plots.long)[i] <- title
  plots.t2[[i]]<- ggplot(data = df, mapping = aes(x = t2.vector, y = long.vector, color = t1.vector)) + geom_point(aes(alpha = 0.3)) + ggtitle(title)
  names(plots.t2)[i] <- title
  plots.t1[[i]]<- ggplot(data = df, mapping = aes(x = t1.vector, y = long.vector, color = t2.vector)) + geom_point(aes(alpha = 0.3)) + ggtitle(title)
  names(plots.t1)[i] <- title
  plots.t1_diff[[i]]<- ggplot(data = df, mapping = aes(x = diff.vector, y = t1.vector)) + geom_point(aes(alpha = 0.3)) + ggtitle(title)
  names(plots.t1_diff)[i] <- title
  #Correlations between centile scores
  t1_t2.cor_test <- cor.test(df$t1.vector.qnorm, df$t2.vector.qnorm)
  t1_long.cor_test <- cor.test(df$t1.vector.qnorm, df$long.vector.qnorm)
  t2_long.cor_test <- cor.test(df$t2.vector.qnorm, df$long.vector.qnorm)
  t1_diff.cor_test <- cor.test(df$t1.vector.qnorm, df$diff.vector.qnorm)
  cor_df$label[i] <- title
  cor_df$t1_t2.cor[i] <- t1_t2.cor_test$estimate
  cor_df$t1_t2.p[i] <- t1_t2.cor_test$p.value
  cor_df$t1_long.cor[i] <- t1_long.cor_test$estimate
  cor_df$t1_long.p[i] <- t1_long.cor_test$p.value
  cor_df$t2_long.cor[i] <- t2_long.cor_test$estimate
  cor_df$t2_long.p[i] <- t2_long.cor_test$p.value
  cor_df$t1_diff.cor[i] <- t1_diff.cor_test$estimate
  cor_df$t1_diff.p[i] <- t1_diff.cor_test$p.value
}

cor_df$t1_t2.p_adj <- p.adjust(cor_df$t1_t2.p, method = "BH")
cor_df$t1_diff.p_adj <- p.adjust(cor_df$t1_diff.p, method = "BH")
cor_df$t1_long.p_adj <- p.adjust(cor_df$t1_long.p, method = "BH")
cor_df$t2_long.p_adj <- p.adjust(cor_df$t2_long.p, method = "BH")


print(paste0("Correlation Coef Range t1-t2 centiles: ", range(cor_df[cor_df$t1_t2.p_adj < 0.05,]$t1_t2.cor)))
print(paste0("Correlation Coef Range t1-(t2-t1 diff) centiles ", range(cor_df[cor_df$t1_diff.p_adj < 0.05,]$t1_diff.cor)))
print(paste0("P val for min corr t1-(t2-t1 diff) centiles: ", cor_df[which(cor_df$t1_diff.cor == min(cor_df[cor_df$t1_diff.p_adj < 0.05,"t1_diff.cor"])),"t1_diff.p_adj"]))
print(paste0("P val for max corr t1-(t2-t1 diff) centiles: ", cor_df[which(cor_df$t1_diff.cor == max(cor_df[cor_df$t1_diff.p_adj < 0.05,"t1_diff.cor"])),"t1_diff.p_adj"]))

print(paste0("Correlation Coef Range t1-long centiles: ", range(cor_df[cor_df$t1_long.p_adj < 0.05,]$t1_long.cor)))
print(paste0("P val for min corr t1-long centiles: ", cor_df[which(cor_df$t1_long.cor == min(cor_df[cor_df$t1_long.p_adj < 0.05,"t1_long.cor"])),"t1_long.p_adj"]))
print(paste0("P val for max corr t1-long centiles: ", cor_df[which(cor_df$t1_long.cor == max(cor_df[cor_df$t1_long.p_adj < 0.05,"t1_long.cor"])),"t1_long.p_adj"]))

print(paste0("Correlation Coef Range t2-long centiles: ", range(cor_df[cor_df$t2_long.p_adj < 0.05,]$t2_long.cor)))
```

## Plots showing centile relationships for total cortical GMV
```{r}
# pheno <- cor_df %>% filter(t1_diff.cor == min(cor_df$t1_diff.cor)) %>% pull(label)
# pheno_var_name <- name_mapping[name_mapping$region_name == pheno, "abbreviation"]

cor.test(df.zscore$smri_vol_cdk_total.t1_centiles, df.zscore$smri_vol_cdk_total.diff.qnorm)
cor.test(full.df$smri_vol_cdk_total.t1_centiles, full.df$smri_vol_cdk_total.diff)

totalgray_t1_diff_qnorm_plot <- ggplot(data = df.zscore, mapping = aes(x = smri_vol_cdk_total.diff.qnorm, y = smri_vol_cdk_total.t1_centiles)) +  
  geom_point(alpha = 0.25, color = "#81888F") + 
  geom_smooth(method = "lm") +
  ggpubr::stat_cor(
    label.x = 0.08,
  label.y = 0.9) +
  theme_classic() +
  labs(x = "T2 Centile (qnorm) - T1 Centile (qnorm)", y = "T1 Centile (qnorm)", title = "Total Cerebral Cortex") +
  theme(plot.title =  element_text(hjust = 0.5),
        text = element_text(size = 24))

totalgray_t1_long_qnorm_plot <- ggplot(data = df.zscore, mapping = aes(x = smri_vol_cdk_total.long_centiles, y = smri_vol_cdk_total.t1_centiles)) +  
  geom_point(alpha = 0.25, color = "#81888F") + 
  geom_smooth(method = "lm") +
  ggpubr::stat_cor(
    label.x = 0.08,
  label.y = 0.9) +
  theme_classic() +
  labs(x = "Longitudinal Centile (qnorm)", y = "T1 Centile (qnorm)", title = "Total Cerebral Cortex") +
  theme(plot.title =  element_text(hjust = 0.5),
        text = element_text(size = 24))

totalgray_t1_diff_longcolor_qnorm_plot <- ggplot(data = df.zscore, 
                                                 mapping = aes(x = smri_vol_cdk_total.diff.qnorm, y = smri_vol_cdk_total.t1_centiles,
                                                                                 color = smri_vol_cdk_total.long_centiles)) +  
  geom_point(alpha = 0.25) + 
  scale_color_gradient2(  low = "#2C4A9A", mid = "#7C6B4F", high = "#D98E04", name = "Longitudinal Centile (qnorm)") +
  geom_smooth(method = "lm") +
  ggpubr::stat_cor(
    label.x = 0.08,
  label.y = 0.9) +
  theme_classic() +
  labs(x = "T2 Centile (qnorm) - T1 Centile (qnorm)", y = "T1 Centile (qnorm)", title = "Total Cerebral Cortex") +
  theme(plot.title =  element_text(hjust = 0.5),
        text = element_text(size = 24),
        legend.position = "bottom")

totalgray_t1_long_diffcolor_qnorm_plot <- ggplot(data = df.zscore, 
                                                 mapping = aes(x = smri_vol_cdk_total.long_centiles, y = smri_vol_cdk_total.t1_centiles,
                                                               color = smri_vol_cdk_total.diff.qnorm)) +  
  scale_color_gradient2(low = "#2C4A9A", mid = "#7C6B4F", high = "#D98E04", name = "T2 Centile (qnorm) - T1 Centile (qnorm)") +
  geom_point(alpha = 0.25) + 
  geom_smooth(method = "lm") +
  ggpubr::stat_cor(
    label.x = 0.08,
  label.y = 0.9) +
  theme_classic() +
  labs(x = "Longitudinal Centile (qnorm)", y = "T1 Centile (qnorm)", title = "Total Cerebral Cortex") +
  theme(plot.title =  element_text(hjust = 0.5),
        text = element_text(size = 24),
        legend.position = "bottom")

#Save the panels
if(save_figures == TRUE){
ggsave(plot = totalgray_t1_diff_qnorm_plot, filename=paste0(revisions_panels_folder,'totalgray_t1_diff_qnorm.svg'), dpi=300, width=16, height=14, device = "svg")
ggsave(plot = totalgray_t1_long_qnorm_plot, filename=paste0(revisions_panels_folder,'totalgray_t1_long_qnorm.svg'), dpi=300, width=16, height=14, device = "svg")
ggsave(plot = totalgray_t1_diff_longcolor_qnorm_plot, filename=paste0(revisions_panels_folder,'totalgray_t1_diff_longcolor_qnorm.svg'), dpi=300, width=16, height=14, device = "svg")
ggsave(plot = totalgray_t1_long_diffcolor_qnorm_plot, filename=paste0(revisions_panels_folder,'totalgray_t1_long_diffcolor_qnorm.svg'), dpi=300, width=16, height=14, device = "svg")
}
```

## Plot Correlations between centile scores on surface maps
```{r}
###plot
#T1-T2
#merge with atlas data
#merge with atlas data only regions that are significant in correlation
plot_data_dkt <- merge(cor_df %>% filter(t1_t2.p_adj < 0.05), ggseg::dk, by = "label")
plot_data_aseg <- merge(cor_df %>% filter(t1_t2.p_adj < 0.05), ggseg::aseg, by = "label")

#Setting manual limits (could match the colorbar limits between different variables)
(t1.t2_cor_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t1_t2.cor)) +
  scale_fill_gradient2(low = "#4878CF", mid = "white", high = "#D65F5F", limits = c(-1, 1)) +
  theme_void()+
  labs(title = "T1-T2 cor"))

(t1.t2_cor_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = t1_t2.cor)) +
  scale_fill_gradient2(low = "white", high = "#D65F5F", limits = c(-1, 1)) +
  theme_void() +
  labs(title = "T1-T2 cor"))

if(save_figures == TRUE){
panel_save(t1.t2_cor_dkt_plot, "t1_t2_dsk", "svg", figure_panels_folder)
panel_save(t1.t2_cor_aseg_plot, "t1_t2_aseg", "svg", figure_panels_folder)
}
#t1-long
#merge with atlas data only regions that are significant in correlation
plot_data_dkt <- merge(cor_df %>% filter(t1_long.p_adj < 0.05), ggseg::dk, by = "label")
plot_data_aseg <- merge(cor_df %>% filter(t1_long.p_adj < 0.05), ggseg::aseg, by = "label")

(t1.long_cor_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t1_long.cor)) +
  scale_fill_gradient2(low = "#4878CF", mid = "white", high = "#D65F5F", limits = c(-1, 1)) +
  theme_void()+
  labs(title = "T1-Long Cor"))

(t1.long_cor_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = t1_long.cor)) +
  scale_fill_gradient2(low = "#4878CF", mid = "white", high = "#D65F5F",limits = c(-1, 1)) +
  theme_void() +
  labs(title = "T1-Long Cor"))

if(save_figures == TRUE){
panel_save(t1.long_cor_dkt_plot, "t1_long_dsk", "svg", figure_panels_folder)
panel_save(t1.long_cor_aseg_plot, "t1_long_aseg", "svg", figure_panels_folder)
}

#t1-diff
#merge with atlas data only regions that are significant in correlation
plot_data_dkt <- merge(cor_df %>% filter(t1_diff.p_adj < 0.05), ggseg::dk, by = "label")
plot_data_aseg <- merge(cor_df %>% filter(t1_diff.p_adj < 0.05), ggseg::aseg, by = "label")

(t1.diff_cor_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t1_diff.cor)) +
  scale_fill_gradient2(low = "#4878CF", mid = "white", high = "#D65F5F", limits = c(-1, 1)) +
  theme_void()+
  labs(title = "T1-Diff Cor"))

(t1.diff_cor_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = t1_diff.cor)) +
  scale_fill_gradient2(low = "#4878CF", mid = "white", high = "#D65F5F",limits = c(-1, 1)) +
  theme_void() +
  labs(title = "T1-Diff Cor"))

if(save_figures == TRUE){
panel_save(t1.diff_cor_dkt_plot, "t1_diff_dsk", "svg", revisions_panels_folder)
panel_save(t1.diff_cor_aseg_plot, "t1_diff_aseg", "svg", revisions_panels_folder)
}

#T2-Long
#merge with atlas data only regions that are significant in correlation
plot_data_dkt <- merge(cor_df %>% filter(t2_long.p_adj < 0.05), ggseg::dk, by = "label")
plot_data_aseg <- merge(cor_df %>% filter(t2_long.p_adj < 0.05), ggseg::aseg, by = "label")

#Setting manual limits (could match the colorbar limits between different variables)
(t2.long_cor_dkt_plot <- ggplot(plot_data_dkt) +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t2_long.cor)) +
  scale_fill_gradient2(low = "#4878CF", mid = "white", high = "#D65F5F", limits = c(-1, 1)) +
  theme_void()+
  labs(title = "T2-Long Cor"))

(t2.long_cor_aseg_plot <- ggplot(plot_data_aseg) +
  geom_brain(atlas = aseg,
             aes(fill = t2_long.cor)) +
  scale_fill_gradient2(low = "#4878CF", mid = "white", high = "#D65F5F",limits = c(-1, 1)) +
  theme_void() +
  labs(title = "T2-Long Cor"))

if(save_figures == TRUE){
panel_save(t2.long_cor_dkt_plot, "t2_long_dsk", "svg", figure_panels_folder)
panel_save(t2.long_cor_aseg_plot, "t2_long_aseg", "svg", figure_panels_folder)
}
```
# CMD-Centile Mahalanobis Distance - on centiles as is (not qnormed)

## Calculate CMD
```{r}
length_names <- length(grep(".t1_centiles", names(full.df)))
#Ensure all names are in the same order
names <- names(full.df)[grep(".t1_centiles", names(full.df))] %>% sub(".t1_centiles","",.)
names.t1 <- paste0(names, ".t1_centiles")
names.t2 <- paste0(names, ".t2_centiles")
names.long <- paste0(names, ".long_centiles")
df_var <- data.frame(matrix(ncol = 4, nrow = dim(full.df)[1]))
colnames(df_var) <- c("t1.variance", "t2.variance", "long.variance", "subject_id")
#CMD Calculation
##T1
t1.center <- colMeans(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".t1_centiles"))))
t1.cov_matrix <- cov(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".t1_centiles"))))
t1.md <- sqrt(mahalanobis(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".t1_centiles"))), t1.center, t1.cov_matrix))
full.df$t1.cmd <- t1.md
##T2
t2.center <- colMeans(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".t2_centiles"))), na.rm = TRUE)
t2.cov_matrix <- cov(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".t2_centiles"))), use = "complete.obs")
t2.md <- sqrt(mahalanobis(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".t2_centiles"))), t2.center, t2.cov_matrix))
full.df$t2.cmd <- t2.md
##long
long.center <- colMeans(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".long_centiles"))), na.rm = TRUE)
long.cov_matrix <- cov(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".long_centiles"))), use = "complete.obs")
long.md <- sqrt(mahalanobis(full.df %>% select(all_of(paste0(phenotypes_for_MC, ".long_centiles"))), long.center, long.cov_matrix))
full.df$long.cmd <- long.md

#remove outliers as defined by >5 sd from the mean
test.df <- full.df %>% filter(t1.cmd < mean(full.df$t1.cmd) + 10*sd(full.df$t1.cmd))

#correlations between CMDs across three centiles
cor.test(test.df$t1.cmd, test.df$t2.cmd)
cor.test(test.df$t1.cmd, test.df$long.cmd)
cor.test(test.df$t2.cmd, test.df$long.cmd)

#Plots for CMD correlations
(t1_t2_cmd <- ggplot(data = test.df, mapping = aes(x = t1.cmd, y = t2.cmd)) + 
  geom_point(alpha = 0.3, color = "gray") + 
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  theme_minimal() + labs (x = "Timepoint 1 CMD", y = "Timepoint 2 CMD") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18)))

(t1_long_cmd <- ggplot(data = test.df, mapping = aes(x = t1.cmd, y = long.cmd)) + 
  geom_point(alpha = 0.3, color = "gray") + 
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  theme_minimal() + labs (x = "Timepoint 1 CMD", y = "Conditional-Longitudinal CMD") +
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18)))


(t2_long_cmd <- ggplot(data = test.df, mapping = aes(x = t2.cmd, y = long.cmd)) + 
  geom_point(alpha = 0.3, color = "gray") + 
  geom_smooth(method = "lm", se = FALSE, color = "magenta") +
  theme_minimal()  + labs (x = "Timepoint 2 CMD", y = "Conditional-Longitudinal CMD")+
  theme(axis.text = element_text(size = 16),
        axis.title = element_text(size = 18)))


if(save_figures == TRUE){
ggsave(plot=t1_t2_cmd, filename=paste0(figure_panels_folder,'t1_t2_cmd_plot.svg'), dpi=300, width=12, height=8)
ggsave(plot=t1_long_cmd, filename=paste0(figure_panels_folder,'t1_long_cmd_plot.svg'), dpi=300, width=12, height=8)
ggsave(plot=t2_long_cmd, filename=paste0(figure_panels_folder,'t2_long_cmd_plot.svg'), dpi=300, width=12, height=8)
}
```

## CMD Correlations with variables of interest
```{r}
cor_mat <- cor(full.df[,c("gestAge", "birth_weight_oz_sum_adj", "bw_pgs", "neonatal_comp_total", "General_p.t1", "General_p.t2", "General_p_change", "t1.cmd", "t2.cmd", "long.cmd")], use = "pairwise.complete.obs")

p_mat <- cor_pmat(full.df[,c("gestAge", "birth_weight_oz_sum_adj", "bw_pgs", "neonatal_comp_total", "General_p.t1", "General_p.t2", "General_p_change", "t1.cmd", "t2.cmd", "long.cmd")], use = "pairwise.complete.obs")

(cmd_vars_corr_plot <- ggcorrplot(
  cor_mat,
  p.mat = p_mat,
  type = "lower",
  insig = "blank",
  lab = TRUE
))

if(save_figures == TRUE){
panel_save(cmd_vars_corr_plot, "cmd-vars-cor-matrix", "svg", extra_panels_folder)
}
```

# Plots for Fig 1
## Choose individuals to plot
sex == M, site == site20
NDAR_INV0HH094Z9   M  site20                        0.9367265                      0.9669537                      0.9848603
 NDAR_INVZ04GJXZA   M  site20                        0.5373170                    0.136125660                    0.161806622
 NDAR_INVY0M8DA4H   M  site20                       0.13186969                      0.4087228                      0.3367802
```{r}
# find three individuals with same sex and site with different t1 centiles
full.df %>% filter(smri_vol_cdk_total.t1_centiles > 0.9 & smri_vol_cdk_total.t2_centiles < smri_vol_cdk_total.t1_centiles & sex == "M" & site.t1 == "site20") %>%
  select(c(src_subject_id, sex, site.t1, smri_vol_cdk_total.long_centiles, smri_vol_cdk_total.t1_centiles, smri_vol_cdk_total.t2_centiles))

full.df %>% filter(smri_vol_cdk_total.t1_centiles < 0.15 & smri_vol_cdk_total.t2_centiles > smri_vol_cdk_total.t1_centiles & sex == "M" & site.t1 == "site20") %>%
  select(c(src_subject_id, sex, site.t1, smri_vol_cdk_total.long_centiles, smri_vol_cdk_total.t1_centiles, smri_vol_cdk_total.t2_centiles))

full.df %>% filter(smri_vol_cdk_total.t1_centiles < 0.6 & smri_vol_cdk_total.t1_centiles > 0.4 & smri_vol_cdk_total.t2_centiles > smri_vol_cdk_total.t1_centiles & sex == "M" & site.t1 == "site20") %>%
  select(c(src_subject_id, sex, site.t1, smri_vol_cdk_total.long_centiles, smri_vol_cdk_total.t1_centiles, smri_vol_cdk_total.t2_centiles))

full.df %>% filter(smri_vol_cdk_total.t1_centiles < 0.6 & smri_vol_cdk_total.t1_centiles > 0.4 & smri_vol_cdk_total.t2_centiles < smri_vol_cdk_total.t1_centiles & sex == "M" & site.t1 == "site20") %>%
  select(c(src_subject_id, sex, site.t1, smri_vol_cdk_total.long_centiles, smri_vol_cdk_total.t1_centiles, smri_vol_cdk_total.t2_centiles))
```

## plot all subjects together for both t1 and t2
```{r}
##specify subjects
subjects <- c("NDAR_INV0HH094Z9" , "NDAR_INVZ04GJXZA", "NDAR_INVY0M8DA4H")
colors <- c("#5B8FD1", "#3AA17E", "#D48672")
subject_df <- full.df %>% filter(src_subject_id %in% subjects) %>% 
  select("smri_vol_cdk_total.t1", "age.t1", "sex", "site.t1", "src_subject_id", "interscan_months_t1t2",
         "smri_vol_cdk_total.t2", "age.t2", "site.t2", 
         "smri_vol_cdk_total.t1_centiles", "smri_vol_cdk_total.t2_centiles",
         "smri_vol_cdk_total.long_centiles", "split_ID") %>%
  mutate(color = colors)


sim.df.t1 <- tibble(sex = "M", site = "site20", age = seq(from = min(t1.sOne_train_df$age), 
                                          to = max(t1.sOne_train_df$age),
                                          length.out = 100))
sim.df.t2 <- tibble(sex = "M", site = "site20", age = seq(from = min(t2.sOne_train_df$age), 
                                          to = max(t2.sOne_train_df$age),
                                          length.out = 100))

### PLOT T1 Centiles ###
model <- t1.best_models_One$smri_vol_cdk_total
t1_cent_plot <- make_centile_fan(model, t1.sOne_train_df, "age", show_points=FALSE, 
                                 color_manual="#5E6673", average_over = FALSE, 
                                 sim_data_list = list(sim = sim.df.t1),
                                 desiredCentiles = c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95),
                                 label_centiles = FALSE) + 
   scale_x_continuous(
   breaks = seq(96, 144, by = 12),
   labels = as.character(seq(8, 12, by = 1)),    
   limits = c(105,133)              
  ) +
  scale_y_continuous(breaks = seq(450000, 800000, by=50000),
                     labels = \(y) {round(y/(10^5), 2)},
                     limits = c(450000, 780000) ) +
  geom_point(data = subject_df, aes(x = age.t1, y = smri_vol_cdk_total.t1), color = colors, shape = 16, size = 5, stroke = 1) +
  geom_text(data = subject_df, aes(x = age.t1, y = smri_vol_cdk_total.t1, 
                                 label = glue("{signif(smri_vol_cdk_total.t1_centiles, digits = 2)*100}%"),
                                 color = colors),
            vjust = -1) +
  scale_color_identity() +
  labs(x = "Age (Years) ", 
       y = expression("Total Gray Matter (" * 10^5 * " " * mm^3 * ")"), 
       title = "") + 
  theme_classic(base_size = 24) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
    axis.line.x = element_line(color = 'lightgrey'),
    axis.line.y = element_blank()
    )

### PLOT T2 Centiles ###
model <- t2.best_models_One$smri_vol_cdk_total
t2_cent_plot <- make_centile_fan(model, t2.sOne_train_df, "age", show_points=FALSE, 
                                 color_manual="#5E6673", average_over = FALSE, 
                                 sim_data_list = list(sim = sim.df.t2),
                                 desiredCentiles = c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)) + 
  scale_x_continuous(
   breaks = seq(120, 168, by = 12),
   labels = as.character(seq(10, 14, by = 1)), 
   limits = c(125,167)
  ) +
  scale_y_continuous(breaks = seq(450000, 800000, by=50000),
                     labels = \(y) {round(y/(10^5), 2)},
                     limits = c(450000, 780000) ) +
  geom_point(data = subject_df, aes(x = age.t2, y = smri_vol_cdk_total.t2), color = colors, shape = 16, size = 5, stroke = 1) +
  geom_text(data = subject_df, aes(x = age.t2, y = smri_vol_cdk_total.t2, 
                                 label = glue("{signif(smri_vol_cdk_total.t2_centiles, digits = 2)*100}%"),
                                 color = colors),
            vjust = -1) +
  scale_color_identity() +
  labs(x = "Age (Years) ", 
       y = expression("Total Gray Matter (" * 10^5 * " " * mm^3 * ")"), 
       title = "") + 
  theme_classic(base_size = 24) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
    axis.line.x = element_line(color = 'lightgrey'),
    axis.line.y = element_blank()
    )


t1_cent_plot
t2_cent_plot

# #Save the panels
if(save_figures == TRUE){
ggsave(plot=t1_cent_plot, filename=paste0(revisions_panels_folder,'t1_centplot.svg'), dpi=300, width=16, height=14, device = "svg")
ggsave(plot=t2_cent_plot, filename=paste0(revisions_panels_folder,'t2_centplot.svg'), dpi=300, width=16, height=14, device = "svg")
}
```

## longitudinal sim for above subjects
```{r}
all_long_cent_plot <- long_centile_sim(t1_model = t1.best_models_One$smri_vol_cdk_total,
                                   t2_model = t2.best_models_One$smri_vol_cdk_total,
                                   long_model = long.best_models_One$smri_vol_cdk_total,
                                   t1_og_df = t1.sOne_train_df,
                                   t2_og_df = t2.sOne_train_df,
                                   long_og_df = long.sOne_train_df,
                                   phenotype = "smri_vol_cdk_total",
                                   phenotype_title = "Total Cortical GMV",
                                   t1_centiles = subject_df$smri_vol_cdk_total.t1_centiles,
                                   sex = subject_df$sex,
                                   site = subject_df$site.t1,
                                   age.t1 = subject_df$age.t1,
                                   interscan_interval = subject_df$interscan_months_t1t2,
                                   plot_colors = subject_df$color,
                                   qnorm_plot = TRUE,
                                   true_long_centile = subject_df$smri_vol_cdk_total.long_centiles
                                   )

all_long_cent_plot

all_long_cent_plot_widelimits <- all_long_cent_plot + ylim(-5, 5)

if(save_figures == TRUE){
ggsave(plot=all_long_cent_plot, filename=paste0(revisions_panels_folder,'fig1_long_cent_plot.svg'), dpi=300, width=10, height=16, device = "svg")
ggsave(plot=all_long_cent_plot_widelimits, filename=paste0(revisions_panels_folder,'fig1_long_cent_plot_widelimits.svg'), dpi=300, width=10, height=16, device = "svg")
}
```

## longitudinal sim for a set of hypothetical subjects
```{r}
supp_all_long_cent_plot <- long_centile_sim(t1_model = t1.best_models_One$smri_vol_cdk_total,
                                   t2_model = t2.best_models_One$smri_vol_cdk_total,
                                   long_model = long.best_models_One$smri_vol_cdk_total,
                                   t1_og_df = t1.sOne_train_df,
                                   t2_og_df = t2.sOne_train_df,
                                   long_og_df = long.sOne_train_df,
                                   phenotype = "smri_vol_cdk_total",
                                   phenotype_title = "Total Cortical GMV",
                                   t1_centiles = c(0.01, 0.1, 0.25, 0.5, 0.75, 0.9, 0.99),
                                   sex = "M",
                                   site = "site16",
                                   age.t1 = 12*9,
                                   interscan_interval = 24,
                                   plot_colors = c(
  "#1F4E79", # deep steel blue
  "#2E6F95", # muted ocean blue
  "#4C8DAE", # slate teal-blue
  "#6BA8A9", # dusty teal
  "#8CBF88", # muted sage green
  "#B5C46A", # olive-lime
  "#D9B44A"  # muted golden amber
),
                                   qnorm_plot = TRUE
                                   )

supp_all_long_cent_plot

supp_all_long_cent_plot_widelimits <- supp_all_long_cent_plot + ylim(-5, 5)


if(save_figures == TRUE){
ggsave(plot=supp_all_long_cent_plot, filename=paste0(revisions_panels_folder,'supfig1_long_cent_plot.svg'), dpi=300, width=10, height=16, device = "svg")
ggsave(plot=supp_all_long_cent_plot_widelimits, filename=paste0(revisions_panels_folder,'supfig1_long_cent_plot_widelimits.svg'), dpi=300, width=10, height=16, device = "svg")
}
```

#Neonatal complications supp table
```{r}
get_counts <- function(x) {
  counts <- table(factor(x, levels = c(0,1,999)), useNA = "ifany")
  as.numeric(counts)
}

neonatal_vars <- c("devhx_14a3_p", "devhx_14b3_p", "devhx_14c3_p", "devhx_14d3_p", "devhx_14e3_p", "devhx_14f3_p", "devhx_14g3_p", "devhx_14h3_p")
abcd_neonatal <- read_csv("/mnt/isilon/bgdlab_processing/Eren/ABCD-braincharts/CSV/process_tables_5.1_2.0/update/abcd_long_afterQC_neonatal_baseline.csv")

#table for counts for each question (before splits)
counts_before <- t(sapply(abcd_neonatal[, neonatal_vars], get_counts))
colnames(counts_before) <- c("0_before", "1_before", "999_before", "NA_before")

#merge with t1 for post-split data Ns
abcd_neonatal_post_split <- merge(
  t1.full_test_df %>% select("src_subject_id", "neonatal_comp_total"),
  abcd_neonatal %>% select(all_of(c("src_subject_id", neonatal_vars))),
  by = "src_subject_id",
  all.x = TRUE
)

##histogram
ggplot(t1.full_test_df, aes(x = neonatal_comp_total)) + 
  geom_bar() +
  theme_minimal()

#table for NAs and "don't know"s for each question (after splits -- more subjects removed)
counts_after <- t(sapply(abcd_neonatal_post_split[, neonatal_vars], get_counts))
colnames(counts_after) <- c("0_after", "1_after", "999_after", "NA_after")

#combine into one table
counts_table <- cbind(counts_before, counts_after)

# neonatal vars as column
#counts_table <- cbind(neonatal_variable = rownames(counts_table), counts_table)

write.csv(counts_table, paste0(tableOne_folder,"neonatal_counts_supp_table.csv"), row.names = TRUE)
```
